---
title: "WEIGHTED EDGES GENE COEXPRESSION NETWORK updated"
date: "08/02/2020"
author: "Hokuto Nakayama"
modified: "Min-Yao Jhu & Steven D. Rowland & Karolina Czarnecki"
---

# This script is modified from Hokuto Nakayama's previously published R script for RNA-Seq gene coexpression network analysis deposited on GitHub (Link: https://github.com/Hokuto-GH/gene-coexpression-network-script).

#load package
```{r}
library(reshape)
```

#load data
```{r}
counts=read.csv ("SOM9_all_Cc_tissue_ID_ex.csv")
str(counts)
dim(counts)
```

#remove 0
```{r}
# counts1 <- counts1[rowSums(counts1) > 0,]
# dim(counts1)
```

#add col/rownames
```{r}
# samples=colnames(counts1)
# genes=rownames(counts1)
```

#check histogram
```{r}
# Y.u=t(log(counts1+(1e-6)))
# par(mfrow=c(2,1))
# hist(as.matrix(counts1), breaks=50, main="normalized counts")
# hist(as.vector(Y.u), breaks=50, main="logarithm of the normalized counts")
# par(mfrow=c(1,1))
```

#normalize data instead of (above) transformation
```{r}
library(RSMod)
norm.counts=data_norm(trans = counts)
t.norm.counts=t(norm.counts)
```

#calcurate corr
```{r}
corr.counts=cor(t.norm.counts, method="pearson")
hist(corr.counts[upper.tri(corr.counts)])
```

#calcurate corr
```{r}
# corr.u=cor(Y.u)
# dim(corr.u)
# corr.s=cor(Y.u, method="pearson")
# hist(corr.s[upper.tri(corr.s)])
```

#fisher's Z transformation
```{r}
n=ncol(counts)      ### NOTE: this is your sample size (library#)
z.s=0.5*log((1+corr.counts)/(1-corr.counts))
summary(z.s[upper.tri(z.s)])
hist(z.s[upper.tri(z.s)])
```

#cut off
```{r}
z.s[is.na(z.s)] <- 0

thre.z=qnorm(0.93)  ## normal quanitle
adjcent.z=abs(z.s)>thre.z  ## symmetric ajacency matrix: 1: there is an edge; 0 : there is no edge
diag(adjcent.z)=0  ## genes do not connect themselves in the network
# rownames(adjcent.z)=rownames(corr.u)
# colnames(adjcent.z)=colnames(corr.u)
sum(adjcent.z, na.rm = TRUE)/2
```

## WEIGHTED ADJACENCY MATRIX
```{r}
z.s[is.na(z.s)] <- 0

dim(z.s)
dim.karo=nrow(z.s)
pos.thresh=qnorm(0.93)
neg.thresh=-qnorm(0.93)
karo.matrix=matrix(nrow=nrow(z.s),ncol = ncol(z.s))

 for(k in 1:dim.karo){
  for(j in (k+1):dim.karo){
    if(j>dim.karo){break}
        
    if(z.s[k,j]>pos.thresh){
    karo.matrix[k,j]=1
      }

    if(z.s[k,j]< neg.thresh){
        karo.matrix[k,j]=-1
          }
  }}

colnames(karo.matrix)=colnames(z.s)
rownames(karo.matrix)=rownames(z.s)

karo.matrix[is.na(karo.matrix)] <- 0
sum(abs(karo.matrix))

```

#save the matrix data
```{r}
# write.csv(adjcent.z, "SOM9_edges.csv")
```


```{r}
## convert to edgelist for use in cytoscape

library(igraph)
g <- graph.adjacency(karo.matrix,weighted = TRUE,diag = FALSE)
dim(karo.matrix)
karo_edge=get.data.frame(g)
write.csv(karo_edge,file="SOM9_ID_ex_Edgelist_0.93.csv")
karo_edge
```

#calculate community
```{r}
index=rowSums(adjcent.z)>0
weight.adjcent.z=adjcent.z[index,index]
dim(adjcent.z)
dim(weight.adjcent.z)
library(igraph)
g.temp=graph.adjacency(weight.adjcent.z, mode="undirected", diag=FALSE)

community.fastgreedy=fastgreedy.community(g.temp)
community.fastgreedy

table(community.fastgreedy$membership)
```

# Betweenness
```{r}
hist(betweenness(g.temp))
b <- betweenness(g.temp, normalized=TRUE)
```

#edge#
```{r}
df.z.g=abs(rowSums(weight.adjcent.z))
hub <- df.z.g
```

#commu#
```{r}
c <- community.fastgreedy$membership
```

#save the data
```{r}
key <- cbind(b, hub, c)
dim(key)
write.csv(data.frame(key),"SOM9_ID_ex_0.93_node_attributes.csv")
```

#visualization
```{r}
V(g.temp)$color <- "gray57"
for(i in 1:length(unique(community.fastgreedy$membership))){
  V(g.temp)[community.fastgreedy$membership==i]$color=i
 
  if(length(which(community.fastgreedy$membership==i))<7){
    V(g.temp)[community.fastgreedy$membership==i]$color="grey"
  }
}
V(g.temp)[df.z.g>25]$color <- "darkred" # You may also want to play around with hub gene thresholds
v.label=rep("",length(V(g.temp)))
v.label=V(g.temp)$name  # if you want to put gene name
v.size=rep(5,length(V(g.temp)))
# v.size[V(g.temp)$name %in% "AP2"]=4 # if you want to change size of specific nodes
V(g.temp)$shape <- "circle"
pdf("SOM9_ID_ex_0.93_Network.pdf", useDingbats=FALSE)
plot(g.temp, layout=layout.kamada.kawai(g.temp), vertex.size=v.size, vertex.frame.color=NA,vertex.label=v.label, vertex.label.cex=0.05,edge.color="gray57", edge.width =0.2)
# dev.off()
#plot(g.temp)
```

```{r}
#write.graph(g.temp,"SOM9.pdf","gml")
```

```{r}
# transfer to cytoscape
library(RCy3)
createNetworkFromIgraph(g,"SOM9_ID_ex_0.93")
```
